---
title: "Homework_2"
author: "Apoorva Srinivasan"
date: "10/2/2018"
output: github_document
---
```{r}
library(tidyverse)
library(ggplot2)
library(readxl)
library(p8105.datasets)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r p1.1}

nyc_transit = read_csv(file = "./hw2_data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") 
  nyc_transit = janitor::clean_names(nyc_transit) %>%
    select(line:entry, vending, ada) %>%
    mutate(entry = recode(entry, "YES" = TRUE, "NO" = FALSE ))
  
   
```

We're given a dataset containing information about nyc transit subway entrance and exit details. After editing out what is not required for this homework, we now have a new dataset with variable that are line, station name, station lattitude, longitude, route 1 to 11, entrance type, entry, vending and compliance with ADA.

Data cleaning steps

* Imported data using read_csv through relative path to ensure reproducability
* Cleaned the variable names using the janitor function which converted all column names into lower snake case.
* Retained the required variables using select to declutter our data.
* Converted the entry variable from character to logical variable using recode.

The dimensions of the data set are  `r nrow(nyc_transit)` x `r ncol(nyc_transit)`. 

Are these data tidy?

No, the data doesn't look tidy because the data for the route variable are spread across eleven columns which makes it difficult to read. 

Further questions:


How many distinct stations are there?
```{r p1.2}
distinct(nyc_transit, line, station_name, .keep_all = TRUE) %>%
  nrow()
```



PROBLEM 2




